---
output: 
  pdf_document:
    citation_package: biblatex
    latex_engine: pdflatex
bibliography: ref.bib
header-includes:
- \usepackage{hyperref}
- \usepackage{array}
- \usepackage{caption}
- \usepackage{graphicx}
- \usepackage{siunitx}
- \usepackage{multirow}
- \usepackage{hhline}
- \usepackage{calc}
- \usepackage{tabularx}
- \usepackage{fontawesome}
- \usepackage[para,online,flushleft]{threeparttable}
- \usepackage{fvextra}
- \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
biblio-style: numeric
title: "Replication code for 'Dynamic Hierarchical Factor Models'"
thanks: "Thanks to Dr. Steven V. Miller for the R Markdown template (http://github.com/svmiller)."
author: 
- name: Anthony M. Thomas, Jr.
  affiliation: The University of Texas at Arlington
abstract: "Insert text."
keywords: "dynamic factor model, gibbs sampling"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin = 0.75in
fontfamily: courier
fontsize: 11pt
# spacing: double
# endnote: no
---

Packages
```{r}
# knitr         for
# markdown      for
# tidyverse     for piping commands with %>% 
# R.matlab      for readMat function
# matlab        for
# mSTEM         for conv function
# signal        for filter function
# fBasics       for vec function

list.pacs <- c("knitr", "markdown", "tidyverse", "R.matlab", "matlab", "mSTEM", "signal", 
               "fBasics")
sapply(list.pacs, require, character.only = TRUE)
renv::snapshot()

# Clear environment
rm(list = ls())
```

```{r setup}
#| include = FALSE
knitr::opts_chunk$set(cache = TRUE, message = FALSE, warning = FALSE,
                      fig.path = '_figs/',
                      cache.path = '_cache/',
                      fig.process = function(x) {
                          x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
                          if (file.rename(x, x2)) x2 else x
                          })
```


# Introduction
Factor modeling has become a widely used dimension reduction technique, but it is often criticized because the factors are difficult to interpret. Interpretation of the factors can be made easier when the data is organized into blocks. 

```{r}
load("data/dhfm_data.Rdata")
```


There are $T = 227$ time series observations on $N = 445$ variables. The data is organized into $B = 5$ blocks. Three of those blocks are organized into sub-blocks. $N_{b,s}$ denotes the number of variables in sub-block $s$ or block $b$. $S_b$ denotes the number of sub-blocks in block $b$, where where $b \in \left\{ 1, 2, \ldots, B \right\}$. $K_F$ denotes the number of common factors. $K_G(b)$ denotes the number of block-level factors in block $b$. $K_H(b, s)$ denotes the number of sub-block-level factors in sub-block $s$ of block $b$, where $s \in \left\{ 1, 2, \ldots, S_b \right\}$.
 
```{r}
bigT <- nrow(data[[1]][[1]])
bigN <- ncol(matrix(unlist(data), nrow = bigT))
B <- 5
S_b <- c(3, 2, 2, 0, 0)
N_bs <- matrix(0, nrow = sum(S_b != 0), ncol = sum(S_b != 0))
for (b in 1:sum(S_b != 0)) {
        if (S_b[b] != 0) {
                for (s in 1:S_b[b]) {
                        N_bs[b, s] <- ncol(data[[b]][[s]])
                }
        }
}
K_F <- 1
K_G <- c(1, 1, 1, 1, 1)
K_H <- vector("list", B)
K_H[[1]] <- c(1, 1, 2)
K_H[[2]] <- c(2, 1)
K_H[[3]] <- c(1, 1)
K_H[[4]] <- K_H[[5]] <- 0

q_F <- 1 # lag order of transition equation
q_G <- rep(1, B)
q_H <- rep(1, B)
q_Z <- rep(1, B)

l_F <- rep(0, B)
l_G <- rep(0, B)
l_H <- rep(0, B)
```

## Preliminary PCA

The `pc_T()` function does PCA on the Sum of Squares and Cross Products matrix. The `do_PC_level4()` function extracts principal components at various levels of the hierarchy.
```{r}
source("code/_functions/pc_T.R")
source("code/_functions/do_PC_level4.R")

tmp <- do_PC_level4(bigZ_ns, S_b, K_F, K_G, K_H)
F_pc <- tmp$F_PC
G_pc <- tmp$G_PC
H_pc <- tmp$H_PC
```

Initialize factors as PC estimates plus random noise.
```{r}

```

Now initialize parameters based on starting values for the factors. First, regress starting values of aggregate factors on their own lags to obtain starting values for the $\Psi_F$ and $\Sigma_F$.

```{r}
trimr <- function (x, a, b) {
 #  Description: Returns matrix/vector stripped of specified rows.
 #  
 #  Inputs:
 #      x : M by N input matrix
 #      a : first a rows to strip
 #      b : last b rows to strip
 #    
 #  Outputs:
 #      z : (M-a) by (N-b) stripped matrix
    
    x <- as.matrix(x)
    m <- nrow(x)
    h1 <- a + 1
    h2 <- m - b
    
    
    return(z)
}

lagn <- function (x, n) {
#  Description: Returns lagged matrix/vector based on n. The rows are shifted 
#               down by the value of n and zeros are filled in the blank rows.
#  
#  Inputs:
#      x  : input matrix (or vector)
#      n  : number of lags
#    
#  Outputs:
#      y : lagged matrix (or vector)
        
        x = matrix(x)
        nt <- nrow(x)
        nc <- ncol(x)
        if (n > 0) {
                x1 <- trimr(x, 0, n)
                y <- rbind(matrix(0, nrow = n, ncol = nc), x1)
        } else if (n < 0) {
                x1 <- trimr(x, abs(n), 0)
                y <- rbind(x1, matrix(0, nrow = abs(n), ncol = nc))
        } else {
            y <- x
        }
        
        return(y)
}

paramsF <- list(B = B, q_G = q_G, K_blocks = K_G, q_F = q_F*matrix(1, K_F, 1), 
                K_facs = K_F, l_F = l_F)



compute_resids <- function (obs, fac, Lambda) {

    fit <- matrix(0, nrow = nrow(obs), ncol = ncol(obs))
    fit <- fit + fac[, 1] %*% t(Lambda[, 1])
    e <- obs - fit
    return(e)
}

switch_rows <- function (X, id) {
        id1 <- id; id2 <- flipud(id)
        temp <- X
        X[id1, ] <- temp[id2, ]
        X[id2, ] <- temp[id1, ]
        return(X)
}

switch_cols <- function (X, id) {
        id1 <- id; id2 <- t(flipud(t(id)))
        temp <- X
        X[, id1] <- temp[, id2]
        X[, id2] <- temp[, id1]
        return(X)
}

startprior <- list(Psi = list(mean = 0, var = 1e10), Sigma = list(shape = 0, dof = 0))

tmp <- draw_psi_sigma(f, 0.5*matrix(1, K_F, 1), matrix(1, K_F, 1), startprior, paramsF, sig_fix)
psi_F <- tmp$PSI; sig2_F <- tmp$SIG

F_lags <- f
lambda_F <- t( solve(t(F_lags) %*% F_lags, t(F_lags) %*% allG) )
Lambda_F <- as.matrix(lambda_F[, 1])

signlam <- NULL
for (i in 1:K_F) {
     signlam[i] <- sign(Lambda_F[i])
     f[, i] <- f[,i] * signlam
     Lambda_F[, i] <- Lambda_F[, i] * signlam[i]
}
Lambda_F[1, 1] <- 1

eG <- eZ <- eH <- vector(mode = "list", length = B)
psi_G <- psi_Z <- vector(mode = "list", length = B)
psi_H <- vector(mode = "list", length = sum(S_b != 0))
sig2_G <- sig2_Z <- sig2_H <- vector(mode = "list", length = B)
Lambda_G <- Lambda_Gsub <- vector(mode = "list", length = B)
Lambda_H <- vector(mode = "list", length = sum(S_b != 0))

for (i in 1:sum(S_b != 0)) {
        Lambda_H[[i]] <- vector(mode = "list", length = S_b[i])
        for (j in 1:S_b[i]) {
                Lambda_H[[i]][[j]] <- matrix(NA, nrow = ncol(data[[i]][[j]]), ncol = K_H[[i]][[j]])
        }
}

for (b in 1:B) {
     lambda_F_mat <- Lambda_F
     lambda_Fb <- as.matrix(lambda_F_mat[b, ])
     eG[[b]] <- compute_resids(g[[b]], f, lambda_Fb)
     paramsF <- list(B = B, q_G = q_G, K_blocks = K_G, q_F = q_F*matrix(1, K_F, 1), 
                K_facs = K_F, l_F = l_F)
     paramsGb <- list(q_F = 1, K_facs = 1, l_F = 0)
     tmp <- draw_psi_sigma(eG[[b]], 0.5*matrix(1, K_G[b], 1), matrix(1, K_G[b], 1), startprior, paramsGb, sig_fix)
     psi_G[[b]] <- tmp$PSI
     sig2_G[[b]] <- tmp$SIG
     
     if (S_b[b] == 0) {
          yy <- data[[b]]
          paramsG <- list(q_G = q_Z[b], K_blocks = matrix(1, ncol(yy), 1), 
                          q_F = q_G[b] * matrix(1, K_G[b], 1), 
                K_facs = K_F, l_F = l_F)
          gb_lags <- g[[b]]
          
          # regress individual variables on subblock factor starting values
          lambda_Gb <- 
          Lambda_G[[b]] <- t(solve( t(gb_lags) %*% gb_lags, t(gb_lags) %*% yy ) )

          # Find Zbi whose loading on Gb is closest to 1 in absolute value
          mindist <- apply(abs(Lambda_G[[b]]) - matrix(1, nrow = nrow(Lambda_G[[b]]), ncol = ncol(Lambda_G[[b]])), 2, function(x) {sort(x, decreasing = T)})
          neworder <- apply(abs(Lambda_G[[b]]) - matrix(1, nrow = nrow(Lambda_G[[b]]), ncol = ncol(Lambda_G[[b]])), 2, function(x) {order(x, decreasing = T)})
          Lambda_G[[b]] <- switch_rows(Lambda_G[[b]], neworder[1, ])
          bigZ[[b]] <- switch_cols(bigZ[[b]], neworder[1, ])
          data[[b]] <- switch_cols(data[[b]], neworder[1, ])
          signlam <- sign(Lambda_G[[b]][1, 1])
          g[[b]] <- g[[b]] * signlam
          Lambda_G[[b]] <- Lambda_G[[b]] * signlam
          Lambda_G[[b]][1, 1] <- 1
          
          eZ[[b]] = compute_resids(yy, g[[b]], Lambda = Lambda_G[[b]])
          paramsZ = list(q_F = q_Z[b] * matrix(1, ncol(yy), 1), 
                         K_facs = ncol(yy), l_F = matrix(0, ncol(yy), 1))
          tmp <- draw_psi_sigma(eZ[[b]], 0.5 * matrix(1, ncol(yy), 1),
                                       matrix(1, ncol(yy), 1), startprior,
                                       paramsZ, 0.5)
          psi_Z[[b]] <- tmp$PSI
          sig2_Z[[b]] <- tmp$SIG
          
     } else { # if S_b[b] != 0)
          # Lambda_G <- matrix(, nrow = bigT, ncol = 0)
          # g[[b]] <- g[[b]]
          gb_lags <- g[[b]]
          Hb <- matrix(unlist(h[[b]]), nrow = bigT)

          # regress individual variables on subblock factor starting values
          lambda_Gb <- t( solve(t(gb_lags) %*% gb_lags, t(gb_lags) %*% Hb) )
          Lambda_Gsub[[b]] <- as.matrix(lambda_Gb[, K_G[b]])
          signlam <- sign(Lambda_Gsub[[b]][1, 1])
          g[[b]][, 1] <- g[[b]] * signlam
          Lambda_Gsub[[b]][, 1] <- as.matrix(Lambda_Gsub[[b]][, 1]) * signlam
          Lambda_Gsub[[b]][1, 1] <- 1

          for (s in 1:S_b[b]) {
               yy <- data[[b]][[s]]
               if (s == 1) {
                    Hbs_indx <- 1:K_H[[b]][s]
               } else {
                    Hbs_indx <- (sum(K_H[[b]][1:(s-1)]) + 1):sum(K_H[[b]][1:s])
               }
               
               lambda_Gsub_mat <- Lambda_Gsub[[b]]
               lambda_Gbs <- lambda_Gsub_mat[Hbs_indx, ]
               eH[[b]][[s]] <- compute_resids(h[[b]][[s]], g[[b]], matrix(lambda_Gbs))
               paramsH <- list(B = N_bs[b, s], q_G = q_Z,
                               K_blocks = matrix(1, ncol(yy), 1),
                               q_F = q_H[b] * matrix(1, K_H[[b]][s], 1),
                               q_G = q_Z[b] * matrix(1, ncol(yy), 1),
                               K_facs = K_H[[b]][s],
                               l_F = l_H[b] * matrix(1, ncol(yy), 1))
               tmp <- draw_psi_sigma(eH[[b]][[s]], 0.5*matrix(1, nrow = K_H[[b]][s], ncol = 1), 
                                     matrix(1, nrow = K_H[[b]][s], ncol = 1), 
                                     startprior, paramsH, 0.5)
               psi_H[[b]][[s]] <- tmp$PSI 
               sig2_H[[b]][[s]] <- tmp$SIG
               
               hbs_lags <- h[[b]][[s]]
               
               lambda_Hbs <- t( solve( t(hbs_lags) %*% hbs_lags, t(hbs_lags) %*% yy ) )
               Lambda_H[[b]][[s]] <- as.matrix(lambda_Hbs[, 1:K_H[[b]][s]])
               
               mindist <- apply(abs(Lambda_H[[b]][[s]]) - matrix(1, nrow(Lambda_H[[b]][[s]]), ncol(Lambda_H[[b]][[s]])), 2, function(x) {sort(x, decreasing = T)})
               neworder <- apply(abs(Lambda_H[[b]][[s]]) - matrix(1, nrow(Lambda_H[[b]][[s]]), ncol(Lambda_H[[b]][[s]])), 2, function(x) {order(x, decreasing = T)})
               
               Lambda_H[[b]][[s]] <- switch_rows(Lambda_H[[b]][[s]], neworder[1, ])
               
               bigZ[[b]][[s]] <- switch_cols(bigZ[[b]][[s]], neworder[1, ])
               
               for (i in 1:K_H[[b]][[s]]) {
                       signlam[i] <- sign(Lambda_H[[b]][[s]][i, i])
                       h[[b]][[s]][ , i] <- h[[b]][[s]][, i] * signlam[i]
                       Lambda_H[[b]][[s]][ , i] <- Lambda_H[[b]][[s]][, i] * signlam[i]
                       }
               
               diag(Lambda_H[[b]][[s]]) <- rep(1, ncol(Lambda_H[[b]][[s]]))
               Lambda_H[[b]][[s]][upper.tri(Lambda_H[[b]][[s]])] <- 0
               
               eZ[[b]][[s]] <- compute_resids(yy, h[[b]][[s]], Lambda_H[[b]][[s]])
               paramsZ <- list(q_F = 1 * matrix(1, ncol(yy), 1), K_facs = ncol(yy), l_F =  matrix(0, ncol(yy), 1))
               tmp <- draw_psi_sigma(eZ[[b]][[s]], 0.5*matrix(1, nrow = N_bs[b, s], ncol = 1), 
                                     matrix(1, nrow = N_bs[b, s], ncol = 1), 
                                     startprior, paramsZ, 0.5)
               psi_Z[[b]][[s]] <- tmp$PSI 
               sig2_Z[[b]][[s]] <- tmp$SIG
               } # end for (s in 1:S_b[b])
        } # end else if (S_b[b] == 0)
} # end for (b in 1:B)
```

Run sampler.
```{r}
jj <- 1

# Set priors
prior <- list("F" = list("Lambda" = list(mean = 0, var = 10), 
                         "Psi" = list(mean = 0.5, var = 10), 
                         "Sigma" = list(shape = 0.01, dof = 4)),
              "G" = list("Lambda" = list(mean = 0, var = 10), 
                         "Psi" = list(mean = 0.5, var = 10), 
                         "Sigma" = list(shape = 0.01, dof = 4)),
              "H" = list("Lambda" = list(mean = 0, var = 10), 
                         "Psi" = list(mean = 0.5, var = 10), 
                         "Sigma" = list(shape = 0.01, dof = 4)),
              "Z" = list("Psi" = list(mean = 0.5, var = 10), 
                         "Sigma" = list(shape = 0.01, dof = 4)))

# Set parameters
#params <- list("B" = B)
n_burn <- 50000
n_keep <- 50000
n_skip <- 50
n_gibbs <- n_burn + n_keep

cal_alpha <- function(FF, Psi, Lambda) {
        
        K_F <- size(FF, 2)
        l_F <- length(Lambda_F) - 1
        q_G <- size(Psi, 2)
        Lambda_lags <- NULL
        for (l in 1:(l_F + 1)) {
                Lambda_lags <- cbind(Lambda_lags, Lambda[[l]])
        }
        l_star <- l_F + q_G
        
        B <- size(Lambda[[1]], 1)
        Lambda_tilde <- NULL
        alpha <- NULL
        for (b in 1:B) {
                lambdab_tilde <- NULL
                for (j in 1:K_F) {
                        indx <- K_F:(K_F*(l_F + 1))
                        lambdabj_tilde <- conv(as.vector(cbind(1, -Psi[b, ])), as.vector(Lambda_lags[b, indx]))
                        lambdab_tilde <- cbind(lambdab_tilde, t(lambdabj_tilde))
                }
                lambdab_tilde <- reshape(t(lambdab_tilde), 1, K_F*(l_star + 1))
                Lambda_tilde <- rbind(Lambda_tilde, lambdab_tilde)
                Flags <- NULL
                for (l in 0:l_star) {
                        Flags <- cbind(Flags, lagn(FF, l))
                }
                alpha[[b]] <- Flags %*% t(lambdab_tilde)
        }
        output <- list(alpha, Lambda_tilde)
        return(output)
}

# for (gibbs in 1:n_gibbs) {
#     # Update global factors and corresponding parameters using last iterations's draws of params and block factors
#     
# }
```
