---
output: 
  pdf_document:
    citation_package: biblatex
    latex_engine: pdflatex
bibliography: ref.bib
header-includes:
- \usepackage{hyperref}
- \usepackage{array}
- \usepackage{caption}
- \usepackage{graphicx}
- \usepackage{siunitx}
- \usepackage{multirow}
- \usepackage{hhline}
- \usepackage{calc}
- \usepackage{tabularx}
- \usepackage{fontawesome}
- \usepackage[para,online,flushleft]{threeparttable}
- \usepackage{fvextra}
- \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
biblio-style: numeric
title: "Replication code for 'Dynamic Hierarchical Factor Models'"
thanks: "Thanks to Dr. Steven V. Miller for the R Markdown template (http://github.com/svmiller)."
author: 
- name: Anthony M. Thomas, Jr.
  affiliation: The University of Texas at Arlington
abstract: "Insert text."
keywords: "dynamic factor model, gibbs sampling"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin = 0.75in
fontfamily: courier
fontsize: 11pt
# spacing: double
# endnote: no
---

Packages
```{r}
# knitr         for
# markdown      for
# tidyverse     for piping commands with %>% 
# R.matlab      for readMat function
# matlab        for
# mSTEM         for conv function
# signal        for filter function
# fBasics       for vec function

list.pacs <- c("knitr", "markdown", "tidyverse", "R.matlab", "matlab", "mSTEM", "signal", 
               "fBasics")
sapply(list.pacs, require, character.only = TRUE)
renv::snapshot()

# Clear environment
rm(list = ls())
```

```{r setup}
#| include = FALSE
knitr::opts_chunk$set(cache = TRUE, message = FALSE, warning = FALSE,
                      fig.path = '_figs/',
                      cache.path = '_cache/',
                      fig.process = function(x) {
                          x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
                          if (file.rename(x, x2)) x2 else x
                          })
```


# Introduction
Factor modeling has become a widely used dimension reduction technique, but it is often criticized because the factors are difficult to interpret. Interpretation of the factors can be made easier when the data is organized into blocks. 

```{r}
load("data/dhfm_data.Rdata")
```

There are $T = 227$ time series observations on $N = 445$ variables. The data is organized into $B = 5$ blocks. Three of those blocks are organized into sub-blocks. $N_{b,s}$ denotes the number of variables in sub-block $s$ or block $b$. $S_b$ denotes the number of sub-blocks in block $b$, where where $b \in \left\{ 1, 2, \ldots, B \right\}$. $K_F$ denotes the number of common factors. $K_G(b)$ denotes the number of block-level factors in block $b$. $K_H(b, s)$ denotes the number of sub-block-level factors in sub-block $s$ of block $b$, where $s \in \left\{ 1, 2, \ldots, S_b \right\}$.
 
```{r}
bigT <- nrow(data[[1]][[1]])
bigN <- ncol(matrix(unlist(data), nrow = bigT))
B <- length(data)
S_b <- c(3, 2, 2, 0, 0)
N_bs <- matrix(0, nrow = sum(S_b != 0), ncol = sum(S_b != 0))
for (b in 1:sum(S_b != 0)) {
        if (S_b[b] != 0) {
                for (s in 1:S_b[b]) {
                        N_bs[b, s] <- ncol(data[[b]][[s]])
                }
        }
}
K_F <- 1
K_G <- c(1, 1, 1, 1, 1)
K_H <- vector("list", B)
K_H[[1]] <- c(1, 1, 2)
K_H[[2]] <- c(2, 1)
K_H[[3]] <- c(1, 1)
K_H[[4]] <- K_H[[5]] <- 0

q_F <- 1 # lag order of transition equation
q_G <- rep(1, B)
q_H <- rep(1, B)
q_Z <- rep(1, B)

l_F <- rep(0, B)
l_G <- rep(0, B)
l_H <- rep(0, B)
```

## Preliminary PCA

I believe that the `nbplog()` function estimates the number of factors. 
```{r}
nbplog <- function(X, kmax, jj, DEMEAN) {
    bigT <- nrow(X)
    bigN <- ncol(X)
    NT <- bigT * bigN
    NT1 <- bigT + bigN
    CT <- matrix(0, nrow = 1, ncol = kmax)
    ii <- 1:kmax
    if (jj == 1){
        CT[1, ] <- log(NT / NT1) * ii * NT1 / NT
    } else if (jj == 2) {
        CT[1, ] <- (NT1 / NT) * log(min(bigN, bigT)) * ii
    } else if (jj == 3) {
        CT[1, ] <- ii * log(min(bigN, bigT)) / min(bigN, bigT)
    } else if (jj == 4) {
        CT[1, ] <- 2 * ii / bigT
    } else if (jj == 5) {
        CT[1, ] <- log(bigT) * ii / bigT
    } else if (jj == 6) {
        CT[1, ] <- 2 * ii * NT1 / NT
    } else if (jj == 7) {
        CT[1, ] <- log(NT) * ii * NT1 / NT
    }
    
    if (DEMEAN == 2) {
        X <- scale(X)
    } else if (DEMEAN == 1) {
        X <- scale(X, scale = FALSE)
    } else {
        X <- X
    }
    
    IC1 <- matrix(0, nrow = nrow(CT), ncol = kmax + 1)
    Sigma <- matrix(0, nrow = 1, ncol = kmax + 1)
    if (bigT < bigN) {
        SVD.XX <- svd(X %*% t(X))
        ev <- SVD.XX$u
        eigval <- SVD.XX$d
        sumeigval <- cumsum(eigval) / sum(eigval)
        Fhat0 <- sqrt(bigT) * ev
        Lambda0 <- t(X) %*% Fhat0 / bigT
    } else {
        SVD.XX <- svd(t(X) %*% X)
        ev <- SVD.XX$u
        eigval <- SVD.XX$d
        sumeigval <- cumsum(eigval) / sum(eigval)
        Lambda0 <- sqrt(bigN) * ev
        Fhat0 <- X %*% Lambda0 / bigN
    }
    
    if (jj < 8) {
        for (i in seq(kmax, 1, by = -1)) {
            Fhat <- Fhat0[, 1:i]
            lambda <- Lambda0[, 1:i]
            chat <- Fhat %*% t(lambda)
            ehat <- X - chat
            Sigma[i] <- mean(base::sum(ehat^2 / bigT))
            IC1[, i] <- log(Sigma[i]) + CT[, i]
        }
        Sigma[kmax + 1] <- mean(base::sum(X^2 / bigT))
        IC1[, kmax + 1] <- log(Sigma[kmax + 1])
        ic1 <- apply(t(IC1), 2, which.min)
        ic1 <- ic1 * as.numeric(ic1 <= kmax)
    }
    
    if (jj == 8) {
        for (j in 1:length(sumeigval)) {
            if (sumeigval[j] >= 0.5) {
                ic1 <- j
            }
        }
    }
    Fhat <- Fhat0[, 1:kmax]
    Lambda <- Lambda0[, 1:kmax]
    chat <- Fhat %*% t(Lambda)
    
    ret.list <- list(ic1 = ic1, chat = chat, Fhat = Fhat, eigval = eigval)
    return(ret.list)
}

# Xb <- data[[1]]
# Xb <- matrix(unlist(data[[1]]), nrow = nrow(Xb[[1]]))
# tmp <- nbplog(Xb, 8, 2, 2)
# ic_Gb <- tmp$ic1
# chat_Gb <- tmp$chat
# GX_b <- tmp$Fhat
# eigval_Gb <- tmp$eigval
```

The `do_PC_level4()` and `pc_T()`functions extracts principal components at various levels of the hierarchy. I don't fully understand how the algorithm works.
```{r}
pc_T <- function (X, nfac) {
    # Principal components with normalization F'F/T = I
    bigT <- nrow(X)
    bigN <- ncol(X)
    XX <- X %*% t(X)
    SVD.XX <- svd(XX)
    U <- SVD.XX$u
    eigval <- diag(SVD.XX$d)
    F.hat <- U[ , 1:nfac] * sqrt(bigT)
    Lambda <- t(X) %*% F.hat
    e.hat <- X - F.hat %*% t(Lambda)
    SS <- diag(eigval)
    ret.list <- list(Lambda = Lambda, F.hat = F.hat, e.hat = e.hat, SS = SS)
    return(ret.list)
    
        # bigt <- nrow(y); bign <- ncol(y)
        # yy <- y %*% t(y)
        # svd.yy <- svd(yy)
        # Fhat0 <- svd.yy$u; eigval <- diag(svd.yy$d); Fhat1 <- svd.yy$v
        # fhat <- Fhat0[, 1:nfac] * sqrt(bigt)
        # lambda <- t(y) %*% fhat / bigt
        # ehat <- y - fhat %*% t(lambda)
        # ss <- diag(eigval)
        # returnList <- list(lambda = lambda, fhat = fhat, ehat = ehat, ss = ss)
}
do_PC_level4 <- function(bigX, S_b, K_max, K_F, K_G, K_H) {
    # bigX is the data 
    # S_b is the vector that has the number of sub-blocks in each block
    
    B <- length(S_b)
    bigT <- nrow(bigX[[1]][[1]])
    bigXmat <- matrix(NA, nrow = bigT, ncol = 0)
    bigGmat <- matrix(NA, nrow = bigT, ncol = 0)
    bigHmat <- matrix(NA, nrow = bigT, ncol = 0)
    G <- GX <- vector("list", B)
    H <- vector("list", sum(S_b != 0))
    
    for (b in 1:B) {
        Xb <- bigX[[b]]
        if (is.list(Xb)) {
            Xb <- matrix(unlist(Xb), nrow = nrow(Xb[[1]]))
        }
        bigXmat <- cbind(bigXmat, Xb)
        GX[[b]] <- pc_T(scale(Xb), K_G[b])$F.hat
        for (s in 1:S_b[b]) {
            if (S_b[b] > 0){
                Zbs <- bigX[[b]][[s]]
                H[[b]][[s]] <- pc_T(scale(Zbs), K_H[[b]][[s]])$F.hat
                bigHmat <- cbind(bigHmat, H[[b]][[s]])
            }
        }
        if (S_b[b] > 0) {
            G[[b]] <- pc_T(scale(matrix(unlist(H[[b]]), nrow = bigT)), K_G[b])$F.hat
        } else {
            G[[b]] <- GX[[b]]
        }
        bigGmat = cbind(bigGmat, G[[b]])
    }
    
    ic_FH <- nbplog(bigHmat, min(K_max, ncol(bigHmat)), 2, 2)$ic1
    FH <- pc_T(bigHmat, ic_FH)$F.hat
    
    FX <- pc_T(scale(bigXmat), K_F)$F.hat
    
    ic_F <- nbplog(bigGmat, min(K_max, ncol(bigGmat)), 2, 2)$ic1
    FF <- pc_T(bigGmat, min(max(ic_F, 1), K_F))$F.hat
    
    ret.list <- list(FF = FF, G = G, H = H, FH = FH, FX = FX, GX = GX)
    return(ret.list)
}

PC_level4 <- do_PC_level4(data, S_b, 8, K_F, K_G, K_H)
F_pc <- PC_level4$FF
G_pc <- PC_level4$G
H_pc <- PC_level4$H
FH_pc <- PC_level4$FH
FX_pc <- PC_level4$FX
GX_pc <- PC_level4$GX
```

Initialize factors as PC estimates plus random noise.
```{r}
sig_fix = .1;
noise_var = .2;

g <- vector(mode = "list", length = B)
h <- vector(mode = "list", length = sum(S_b != 0))

for (b in 1:B) {
        g[[b]] <- matrix(G_pc[[b]][1:K_G[b]] + noise_var*rnorm(bigT), nrow = bigT)
        if (S_b[b] == 0) {
                data[[b]] <- scale(data[[b]])
                for (k in 1:K_G[b]) {
                        g[[b]][, k] <- sign(cor(g[[b]][, k], data[[b]][, k])) * g[[b]][, k]
                }
        } else {
                for (s in 1:S_b[b]) {
                        data[[b]][[s]] <- scale(data[[b]][[s]])
                        h[[b]][[s]] <- matrix(H_pc[[b]][[s]] + noise_var*rnorm(bigT), nrow = bigT)
                        for (k in 1:K_H[[b]][s]) {
                                h[[b]][[s]][, k] <- sign(cor(h[[b]][[s]][, k], data[[b]][[s]][, k])) * h[[b]][[s]][, k]
                        }
                }
                allH <- matrix(unlist(h[[b]]), nrow = bigT)
                for (k in 1:K_G[b]) {
                        g[[b]][, k] <- sign(cor(g[[b]][, k], allH[, k])) * g[[b]][, k]
                }
        }
}

allG <- matrix(unlist(g), nrow = bigT)
f <- matrix(pc_T(allG, K_F)$F.hat, bigT) + noise_var*rnorm(bigT)

for (k in 1:K_F) {
        f[, k] <- sign(cor(f[, k], allG[, k])) * f[, k]
}
```

```{r}
trimr <- function (X, a, b) {
 #  Description: Returns matrix/vector stripped of specified rows.
 #  
 #  Inputs:
 #      X  : input matrix (or vector)
 #      m1 : first a rows to strip
 #      m2 : last b rows to strip
 #    
 #  Outputs:
 #      Z : Stripped matrix (or vector)
    m <- nrow(X)
    h1 <- a + 1
    h2 <- m - b
    Z <- X[h1:h2, ]
    
    return(Z)
}

lagn <- function (x, n) {
#  Description: Returns lagged matrix/vector based on n. The rows are shifted down by the value of n and zeros are filled in the blank rows.
#  
#  Inputs:
#      x  : input matrix (or vector)
#      n  : number of lags
#    
#  Outputs:
#      y : lagged matrix (or vector)
    
        nt <- nrow(x)
        nc <- ncol(x)
        if (n > 0) {
                x1 <- trimr(x, 0, n)
                y <- rbind(matrix(0, nrow = n, ncol = nc), x1)
        }
        if (n < 0) {
                x1 <- trimr(x, abs(n), 0)
                y <- rbind(x1, matrix(0, nrow = abs(n), ncol = nc))
        }
        if (n == 0) {
                y <- x
        }
        return(y)
}

draw_psi_sigma <- function (Ft, psi_F, sig2_F, prior.psi.mean, prior.psi.var, prior.Sigma.shape, prior.Sigma.dof, l_F, q_F, var_fix) {
        K_F <- ncol(Ft)
        max_q_F <- max(q_F)
        max_l_F <- max(l_F)
        bigT <- nrow(Ft)
        PSI <- psi_F; SIG <- sig2_F
        
        for (k in 1:K_F) {
                q_Fk <- q_F[k]
                xpsi_F1 = matrix(0, nrow = max_q_F, 1)
                if (q_Fk > 0) {
                        ystar <- as.matrix(Ft[, k])
                        xstar <- matrix(NA, nrow = bigT, ncol = 0)
                        for (i in 1:q_Fk) {
                                xstar <- cbind(xstar, lagn(f, i))
                        }
                        ystar <- trimr(ystar, max(max_q_F, max_l_F), 0)
                        xstar <- trimr(xstar, max(max_q_F, max_l_F), 0)
                        TT <- nrow(ystar)
                        
                        prior_mean <- prior.psi.mean * matrix(1, nrow = q_Fk, ncol = 1)
                        inv_prior_var <- as.numeric(solve(prior.psi.var)) * diag(1, nrow = q_Fk)
                        sig_inv <- 1/sig2_F[k]
                        
                        post_var <- solve(inv_prior_var + sig_inv * t(xstar) %*% xstar)
                        post_mean <- post_var %*% (inv_prior_var %*% prior_mean + sig_inv %*% t(xstar) %*% ystar)
                        
                        C <- chol(sig2_F[k] * post_var)
                        accept <- 0; counter <- 0
                        while (accept == 0 & counter < 1e3) {
                                psi_F1 <- post_mean + t(C) * rnorm(q_Fk, 1)
                                a <- rbind(1, -psi_F1); ceof <- a[nrow(a):1, ]
                                root <- polyroot(ceof)
                                rootmod <- abs(root)
                                if (min(rootmod) > 1.0001) {
                                        accept = 1
                                } else {
                                        counter <- counter + 1
                                        accept <- 0
                                }
                        }
                        # if (!accept & counter >= 1000) {
                        #      print(counter)
                        # }
                        SSE <- t(ystar - xstar %*% psi_F1) %*% (ystar - xstar %*% psi_F1)
                } else {
                        TT <- nrow(Ft)
                        ystar <- as.matrix(Ft[, k])
                        SSE <- t(ystar) %*% ystar
                }
                if (!is.null(var_fix)) {
                        sig2_F1 <- var_fix
                } else {
                        d <- prior.Sigma.shape + SSE
                        c <- rchisq(1, df = TT + prior.Sigma.dof)
                        sig2_F1 <- d/c
                }
                xpsi_F1[1:q_Fk] <- psi_F1[1:q_Fk]
                if (k == 1) {
                        SIG <- sig2_F1
                        PSI <- xpsi_F1
                } else {
                        SIG <- rbind(SIG, sig2_F1)
                        PSI <- rbind(PSI, t(xpsi_F1))
                }
        }
        output <- list(SIG = SIG, PSI = PSI)
        return(output)
}

```

Initialize parameters based on starting values for the factors. First regress starting values of aggregate factors on their own lags to obtain starting values for the psi_F's and sig2_F's.

```{r}
startprior <- list(Psi = list(mean = 0, var = 1e10), Sigma = list(shape = 0, dof = 0))
args <- list(Ft = f, psi_F = 0.5*matrix(1, nrow = K_F, ncol = 1), 
             sig2_F = matrix(1, nrow = K_F, ncol = 1), 
             prior.psi.mean = startprior$Psi$mean, prior.psi.var = startprior$Psi$var, 
             prior.Sigma.shape = startprior$Sigma$shape, prior.Sigma.dof = startprior$Sigma$dof, 
             l_F = l_F, q_F = q_F, var_fix = sig_fix)
psi_F <- do.call(draw_psi_sigma, args)$PSI; sig2_F <- do.call(draw_psi_sigma, args)$SIG

F_lags <- f
lambda_F <- t( solve(t(F_lags) %*% F_lags, t(F_lags) %*% allG) )
Lambda_F <- as.matrix(lambda_F[, 1])

signlam <- NULL
for (i in 1:K_F) {
     signlam[i] <- sign(Lambda_F[i])
     f[, i] <- f[,i] * signlam
     Lambda_F[, i] <- Lambda_F[, i] * signlam[i]
}
Lambda_F[1, 1] <- 1

eG <- vector(mode = "list", length = B)
eZ <- vector(mode = "list", length = B)
eH <- vector(mode = "list", length = B)
psi_G <-vector(mode = "list", length = B)
psi_Z <- vector(mode = "list", length = B)
psi_H <- vector(mode = "list", length = sum(S_b != 0))
sig2_G <- vector(mode = "list", length = B)
sig2_Z <- vector(mode = "list", length = B)
sig2_H <- vector(mode = "list", length = B)
Lambda_G <- vector(mode = "list", length = B)
Lambda_Gsub <- vector(mode = "list", length = B)

Lambda_H <- vector(mode = "list", length = sum(S_b != 0))
for (i in 1:sum(S_b != 0)) {
        Lambda_H[[i]] <- vector(mode = "list", length = S_b[i])
        for (j in 1:S_b[i]) {
                Lambda_H[[i]][[j]] <- matrix(NA, nrow = ncol(data[[i]][[j]]), ncol = K_H[[i]][[j]])
        }
}




for (b in 1:B) {
     lambda_F_mat <- Lambda_F
     lambda_Fb <- as.matrix(lambda_F_mat[b, ])
     eG[[b]] <- compute_resids(g[[b]], f, lambda_Fb, K_F = K_G[b], l_F = l_G[b])
     args.G <- list(Ft = eG[[b]], psi_F = 0.5*matrix(1, nrow = K_G[b], ncol = 1), 
                  sig2_F = matrix(1, nrow = K_G[b], ncol = 1), 
                  prior.psi.mean = startprior$Psi$mean, prior.psi.var = startprior$Psi$var, 
                  prior.Sigma.shape = startprior$Sigma$shape, prior.Sigma.dof = startprior$Sigma$dof, 
                  l_F = l_G[b], q_F = q_G[b]*rep(1, K_G[b]), var_fix = sig_fix)
     psi_G[[b]] <- do.call(draw_psi_sigma, args.G)$PSI
     sig2_G[[b]] <- do.call(draw_psi_sigma, args.G)$SIG
     
     if (S_b[b] == 0) {
          yy <- data[[b]]
          
          # regress all Z_bn on G_b in order to get starting values for Lambda_Gb
          gb_lags <- g[[b]]
          
          # regress individual variables on subblock factor starting values
          lambda_Gb <- t(solve( t(gb_lags) %*% gb_lags, t(gb_lags) %*% yy ) )
          Lambda_G[[b]] <- as.matrix(lambda_Gb[, 1])

          # Find Zbi whose loading on Gb is closest to 1 in absolute value
          mindist <- apply(abs(Lambda_G[[b]]) - matrix(1, nrow = nrow(Lambda_G[[b]]), ncol = ncol(Lambda_G[[b]])), 2, function(x) {sort(x, decreasing = T)})
          neworder <- apply(abs(Lambda_G[[b]]) - matrix(1, nrow = nrow(Lambda_G[[b]]), ncol = ncol(Lambda_G[[b]])), 2, function(x) {order(x, decreasing = T)})
          Lambda_G[[b]] <- switch_rows(Lambda_G[[b]], neworder[1, ])
          data[[b]] <- switch_cols(data[[b]], neworder[1, ])
          data[[b]] <- switch_cols(data[[b]], neworder[1, ])
          signlam <- sign(Lambda_G[[b]][1, 1])
          g[[b]][, 1] <- g[[b]] * signlam
          Lambda_G[[b]][, 1] <- Lambda_G[[b]][, 1] * signlam
          Lambda_G[[b]][1, 1] <- 1
          
          eZ[[b]] = compute_resids(yy, g[[b]], Lambda = Lambda_G[[b]], K_F = K_G[b], l_F = l_G[b]*matrix(1, nrow = nrow(yy), ncol = 1))
          args.Z <- list(Ft = eZ[[b]], psi_F = 0.5*matrix(1, nrow = K_G[b], ncol = 1), sig2_F = matrix(1, nrow = nrow(yy), ncol = 1), 
                       prior.psi.mean = startprior$Psi$mean, prior.psi.var = startprior$Psi$var, 
                       prior.Sigma.shape = startprior$Sigma$shape, prior.Sigma.dof = startprior$Sigma$dof, 
                       l_F = l_F[b] * matrix(1, nrow = nrow(yy), ncol = 1), q_F = q_Z[b]*matrix(1, nrow = nrow(yy), ncol = 1), var_fix = sig_fix)
          psi_Z[[b]] <- do.call(draw_psi_sigma, args.Z)$PSI; sig2_Z[[b]] <- do.call(draw_psi_sigma, args.Z)$SIG
          
     } else { # if S_b[b] != 0)
          # Lambda_G <- matrix(, nrow = bigT, ncol = 0)
          # g[[b]] <- g[[b]]
          gb_lags <- g[[b]]
          Hb <- matrix(unlist(h[[b]]), nrow = bigT)

          # regress individual variables on subblock factor starting values
          lambda_Gb <- t( solve(t(gb_lags) %*% gb_lags, t(gb_lags) %*% Hb) )
          Lambda_Gsub[[b]] <- as.matrix(lambda_Gb[, K_G[b]])
          signlam <- sign(Lambda_Gsub[[b]][1, 1])
          g[[b]][, 1] <- g[[b]] * signlam
          Lambda_Gsub[[b]][, 1] <- as.matrix(Lambda_Gsub[[b]][, 1]) * signlam
          Lambda_Gsub[[b]][1, 1] <- 1

          for (s in 1:S_b[b]) {
               yy <- data[[b]][[s]]
               if (s == 1) {
                    Hbs_indx <- 1:K_H[[b]][s]
               } else {
                    Hbs_indx <- (sum(K_H[[b]][1:(s-1)]) + 1):sum(K_H[[b]][1:s])
               }
               
               lambda_Gsub_mat <- Lambda_Gsub[[b]]
               lambda_Gbs <- lambda_Gsub_mat[Hbs_indx, ]
               eH[[b]][[s]] <- compute_resids(h[[b]][[s]], g[[b]], lambda_Gbs, K_F = K_H[[b]][[s]], l_F = l_H[b]*ones(nrow(h[[b]][[s]]), 1))
               args.H <- list(Ft = eH[[b]][[s]], psi_F = 0.5*matrix(1, nrow = K_H[[b]][[s]], ncol = 1), 
                              sig2_F = matrix(1, nrow = K_H[[b]][[s]], ncol = 1), 
                              prior.psi.mean = startprior$Psi$mean, prior.psi.var = startprior$Psi$var, 
                              prior.Sigma.shape = startprior$Sigma$shape, prior.Sigma.dof = startprior$Sigma$dof, 
                              l_F = l_H[b]*ones(nrow(h[[b]][[s]]), 1), q_F = q_Z[b]*ones(nrow(yy), 1), var_fix = sig_fix)
               psi_H[[b]][[s]] <- do.call(draw_psi_sigma, args.H)$PSI; sig2_H[[b]][[s]] <- do.call(draw_psi_sigma, args.H)$SIG
               
               hbs_lags <- h[[b]][[s]]
               
               lambda_Hbs <- t( solve( t(hbs_lags) %*% hbs_lags, t(hbs_lags) %*% yy ) )
               Lambda_H[[b]][[s]] <- as.matrix(lambda_Hbs[, 1:K_H[[b]][s]])
               
               mindist <- apply(abs(Lambda_H[[b]][[s]]) - ones(nrow(Lambda_H[[b]][[s]]), ncol(Lambda_H[[b]][[s]])), 2, function(x) {sort(x, decreasing = T)})
               neworder <- apply(abs(Lambda_H[[b]][[s]]) - ones(nrow(Lambda_H[[b]][[s]]), ncol(Lambda_H[[b]][[s]])), 2, function(x) {order(x, decreasing = T)})
               
               Lambda_H[[b]][[s]] <- switch_rows(Lambda_H[[b]][[s]], neworder[1, ])
               
               data[[b]][[s]] <- switch_cols(data[[b]][[s]], neworder[1, ])
               data[[b]][[s]] <- switch_cols(data[[b]][[s]], neworder[1, ])
               
               for (i in 1:K_H[[b]][[s]]) {
                       signlam[i] <- sign(Lambda_H[[b]][[s]][i, i])
                       h[[b]][[s]][ , i] <- h[[b]][[s]][, i] * signlam[i]
                       Lambda_H[[b]][[s]][ , i] <- Lambda_H[[b]][[s]][, i] * signlam[i]
                       }
               
               diag(Lambda_H[[b]][[s]]) <- rep(1, ncol(Lambda_H[[b]][[s]]))
               Lambda_H[[b]][[s]][upper.tri(Lambda_H[[b]][[s]])] <- 0
               
               eZ[[b]][[s]] <- compute_resids(yy, h[[b]][[s]], Lambda_H[[b]][[s]], K_F = K_H[[b]][[s]], l_F = l_H[b]*ones(nrow(h[[b]][[s]]), 1))
               args.Z <- list(Ft = eZ[[b]][[s]], psi_F = 0.5*ones(Nsub[b, s], 1), sig2_F = ones(Nsub[b, s], 1), 
                              prior.psi.mean = startprior$Psi$mean, prior.psi.var = startprior$Psi$var, 
                              prior.Sigma.shape = startprior$Sigma$shape, prior.Sigma.dof = startprior$Sigma$dof, 
                              l_F = l_H[b]*ones(nrow(yy), 1), q_F = q_Z[b]*ones(nrow(yy), 1), var_fix = sig_fix)
               psi_Z[[b]][[s]] <- do.call(draw_psi_sigma, args.Z)$PSI; sig2_Z[[b]][[s]] <- do.call(draw_psi_sigma, args.Z)$SIG
               } # end for (s in 1:S_b[b])
        } # end else if (S_b[b] == 0)
} # end for (b in 1:B)
```

```{r}
#| eval = FALSE, include = FALSE

## Functions ----



switch_rows <- function (X, id) {
        id1 <- id; id2 <- flipud(id)
        temp <- X
        X[id1, ] <- temp[id2, ]
        X[id2, ] <- temp[id1, ]
        return(X)
}
switch_cols <- function (X, id) {
        id1 <- id; id2 <- t(flipud(t(id)))
        temp <- X
        X[, id1] <- temp[, id2]
        X[, id2] <- temp[, id1]
        return(X)
}

# Initialize Sampler ----












# Run Sampler ----

jj <- 1

## Priors ----
prior.F <- list(Lambda.mean = 0, Lambda.var = 10, Psi.mean = 0.5, Psi.var = 10, Sigma.shape = 0.01, Sigma.dof = 4)
prior.G <- list(Lambda.mean = 0, Lambda.var = 10, Psi.mean = 0.5, Psi.var = 10, Sigma.shape = 0.01, Sigma.dof = 4)
prior.H <- list(Lambda.mean = 0, Lambda.var = 10, Psi.mean = 0.5, Psi.var = 10, Sigma.shape = 0.01, Sigma.dof = 4)
prior.Z <- list(Psi.mean = 0.5, Psi.var = 10, Sigma.shape = 0.01, Sigma.dof = 4)
priors <- list(prior.F, prior.G, prior.H, prior.Z)

cal_alpha <- function(FF, Psi, Lambda) {
        
        K_F <- size(FF, 2)
        l_F <- length(Lambda_F) - 1
        q_G <- size(Psi, 2)
        Lambda_lags <- NULL
        for (l in 1:(l_F + 1)) {
                Lambda_lags <- cbind(Lambda_lags, Lambda[[l]])
        }
        l_star <- l_F + q_G
        
        B <- size(Lambda[[1]], 1)
        Lambda_tilde <- NULL
        alpha <- NULL
        for (b in 1:B) {
                lambdab_tilde <- NULL
                for (j in 1:K_F) {
                        indx <- K_F:(K_F*(l_F + 1))
                        lambdabj_tilde <- conv(as.vector(cbind(1, -Psi[b, ])), as.vector(Lambda_lags[b, indx]))
                        lambdab_tilde <- cbind(lambdab_tilde, t(lambdabj_tilde))
                }
                lambdab_tilde <- reshape(t(lambdab_tilde), 1, K_F*(l_star + 1))
                Lambda_tilde <- rbind(Lambda_tilde, lambdab_tilde)
                Flags <- NULL
                for (l in 0:l_star) {
                        Flags <- cbind(Flags, lagn(FF, l))
                }
                alpha[[b]] <- Flags %*% t(lambdab_tilde)
        }
        output <- list(alpha, Lambda_tilde)
        return(output)
}

sample_facs <- function(g, alpha, lambda_tilde, psi_F, sig2_F, psi_G, sig2_G, params) {
        
        B <- params[["B"]]
        q_G <- params[["q_G"]]
        max_q_G <- max(q_G)
        K_G <- params[["K_G"]]
        K_F <- params[["K_facs"]]
        q_F <- params[["q_F"]]
        max_q_F <- max(q_F)
        
        Gmat <- NULL
        sig2_G_vec <- NULL
        psi_G_mat <- NULL
        
        for (b in 1:B) {
                Gmat <- cbind(Gmat, g[[b]])
                sig2_G_vec <- rbind(sig2_G_vec, sig2_G[[b]])
                
                if (max_q_G > 0) {
                        psi_G_mat <- rbind(psi_G_mat, cbind(psi_G[[b]], matrix(0, nrow = K_G[b], ncol = max_q_G - q_G[b])))
                } else {
                        psi_G_mat <- rbind(psi_G_mat, matrix(0, nrow = K_G[b], ncol = 1))
                }
        }
        TT <- size(Gmat, 1); N <- size(Gmat, 2)
        ystar <- matrix(0, nrow = TT, ncol = N)
        for (i in 1:N) {
                require(signal)
                ystar[ , i] <- filter(as.vector(cbind(1, -Psi[b, ])), 1, Gmat[ , i])
        }
        ystar <- t(ystar)
        
        # transition matrix of the filtered data is r = p + 1 + q + 1
        ndim <- size(lambda_tilde, 2)
        psi_F_mat <- zeros(K_F, max_q_F * K_F)
        for (r in 1:K_F) {
                indx <- seq(r, K_F*max_q_F, K_F)
                psi_F_mat[r, indx] <- psi_F[r, ]
        }
        
        A <- zeros(ndim)
        A[1:size(psi_F_mat, 1), 1:(ndim - K_F)] <- eye(ndim - K_F)
        H <- lambda_tilde
        QQ <- zeros(ndim, ndim)
        QQ[1:K_F, 1:K_F] <- diag(sig2_F)
        R <- diag(sig2_G_vec)
        Alpha <- cbind(alpha, zeros(nrows(alpha), ndim - ncol(alpha)))
        
        # Initialize Kalman Filter
        Gtt <- zeros(ndim, 1)
        P00 <- solve(eye(ndim^2) - kronecker(A, A)) %*% vec(QQ)
        Ptt <- reshape(P00, ndim, ndim)
        
        # Kalman filter recursion
        Fmat <- zeros(ndim, TT)
        Pmat <- vector(mode = "list", length = TT)
        
        for (t in 1:TT) {
                Gtt1    <- t(Alpha[t, ]) + A %*% Gtt1           # x(t|t-1)
                Ptt1    <- A %*% Ptt %*% t(A) + QQ              # P(t|t-1)
                ett1    <- ystar[ , t] - H %*% Gtt1             # eta(t|t-1)=y(t)- y(t|t-1)= y(t)-H*x(t|t-1)
                v_ett1  <- H %*% Ptt11 %*% t(H) + R             # var(eta(t|t-1))
                k_gn    <- Ptt1 %*% t(H) %*% solve(v_ett1)      # K=P(t|t-1)H'/ f(t|t-1)
                Gtt     <- Gtt1 + k_gn %*% ett1                 # x(t|t)=x(t|t-1)+ K eta(t|t-1)
                Ptt     <- (eye(ndim) - k_gn %*% H) %*% Ptt1    # P(t|t)= P(t|t-1)-K*H*P(t|t-1)
                
                Fmat[ , t] <- Gtt
                Pmat[[t]] <- Ptt
        }
        
        # Carter-Kohn backward sampling algorithm
        FT_mat <- zeros(TT, K_F)
        G <- t(chol(Pmat[[TT]]))
        FT <- Fmat[ , TT] + G %*% matrix(rnorm(ndim), nrow = ndim, ncol = 1)
        FT_mat[TT, 1:K_F] <- FT[1:K_F, 1]
        jj <- 1:K_F
        
        for (t in (TT-1):1) {
                etT <- FT[jj] - t(Alpha[t + 1, jj]) - A[jj, ] %*% Fmat[ , t]
                v_etT <- A[jj, ] %*% Pmat[[t]] %*%t(A[jj, ]) + QQ[jj, jj]
                k_gn0 <- Pmat[[t]] %*% t(A[jj, ]) %*% solve(v_etT)
                Fmat[ , t] <- Fmat[ , t] + k_gn0 %*% etT
                Pmat[[t]] <- (eye(ndim) - k_gn0 %*% A[jj, ]) %*% Pmat[[t]]
                G <- chol(Pmat[[t]])
                FT <- Fmat[ , t] + G %*% rnorm(ndim)
                FT_mat[t, 1:K_F] <- FT[1:KF, 1]
        }
        done <- 1
        return(FF_mat)
}

draw_lambda <- function(G, FF, psi_G, sig2_G, prior, params) {
      
       B <- params[["B"]]
       q_G <- params[["q_G"]]
       max_q_G <- max(q_G)
       K_G <- params[["K_blocks"]]
       l_F <- params[["l_F"]]
       max_l_F <- max(l_F)
       Lambda_F <- NULL
       count <- 0
       LAMBDA <- vector(mode = "list", length = max_l_F + 1)
       
       for (b in 1:B) {
               Gb <- G[[b]]
               TT <- size(Gb, 1)
               k_G <- size(Gb, 2)
               l_Fb <- l_F[b]
               
               for (i in 1:k_G) {
                       count <- count + 1
                       
                       F_filt <- filter(cbind(1, -psi_G[[b]][i, ]), 1, FF)
                       ystar <- filter(cbind(1, -psi_G[[b]][i, ]), 1, Gb[ , i])
                       Fstar <- F_filt
                       for (j in 1:l_Fb) {
                               Fstar <- cbind(Fstar, lagn(F_filt, j))
                       }
                       Fstar <- trimr(Fstar, max(max_q_G, l_Fb), 0)
                       ystar <- trimr(ystar, max(max_q_G, l_Fb), 0)
                       
                       # bring to right dimension
                       prior_mean <- prior[["mean"]] %*% ones(K_F*(l_Fb + 1), 1)
                       prior_var <- prior[["var"]] %*% eye(K_F*(l_Fb + 1))
                       sig_inv <- 1/sig2_G[[b]][i]
                       
                       # posterior mean and variance
                       post_var <- solve( solve(prior_var) + sig_inv %*% t(Fstar) %*% Fstar )
                       post_mean <- post_var %*% ( solve(prior_var) %*% prior_mean + sig_inv %*% t(Fstar) %*% ystar )
                       
                       # draw from multivariate normal with posterior mean and variance
                       C <- chol(post_var)
                       lambda_Fi <- post_mean + t(C) %*% matrix(rnorm(K_F * (l_Fb + 1)), ncol = 1)
                       Lambda_F <- rbind(Lambda_F, t(lambda_Fi))
                       
               } # end for (i in 1:k_G)
       } # end for (b in 1:B)
       
       # Impose lower triangularity with ones on the diagonal for identification
       
       for (l in 1:(max_l_F + 1)) {
              LAMBDA[[l]] <- Lambda_F[ , ((l-1)*K_F + 1):(l*K_F)]
              for (j in 1:K_F) {
                if (l == 1) {
                  LAMBDA[[l]][j, j] <- 1
                }
                LAMBDA[[l]][j, (j+1):K_F] <- 0
              }
       }
}

Lambda_F <- list(Lambda_F)
for (gibbs in 1:n_gibbs) {
        # Update global factors and corresponding parameters using last iteration's draws of params and block factors
        alpha_F <- cal_alpha(f, matrix(unlist(psi_G), ncol = 1), Lambda_F)$alpha
        lambda_F_tilde <- cal_alpha(f, unlist(psi_G), Lambda_F)$Lambda_tilde
        
        paramsF <- list(B <- B, q_G <- q_G, q_F <- q_F, K_G <- K_G, K_F <- K_facs)
        f <- do.call(sample_facs, c(g, zeros(size(f)), lambda_F_tilde, psi_F, sig2_F, psi_G, sig2_G, paramsF))
} # end for (gibbs in 1:n_gibbs)
```

```{r}
n_burn <- 50000
n_keep <- 50000
n_skip <- 50
n_gibbs <- n_burn + n_keep
```




